---
title: "chapter6.Rmd"
output: html_document
---
Ville Turppo, week 6



## 1) RATS data analysis swapped.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

RATSL <- read.table ("C:/Users/ville/OneDrive/Työpöytä/Open science kurssi 2020/IODS-project/Data/RATSL.csv")
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)
dim(RATSL); head(RATSL); str(RATSL)
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```
It seems that rats in group 1 are lighter than rats in the groups 2 and 3. In groups 2 and 3 the rats are heavier and gain more weight.

Standardized Weight
```{r}
RaS <- RATSL %>% group_by(Time) %>% mutate(stdweight = (Weight - mean(Weight))/sd(Weight) ) %>% ungroup()
glimpse(RaS)
ggplot(RaS, aes(x = Time, y = stdweight, group = ID)) + geom_line(aes(linetype = Group)) + scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) + scale_y_continuous(name = "Weight (standardized)") + theme(legend.position = "top")
```


Summary graphs

```{r}
n <- RaS$Time %>% unique() %>% length()
RaSu <- RaS %>% group_by(Group, Time) %>% summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>% ungroup()
ggplot(RaSu, aes(x = Time, y = mean, linetype = Group, shape = Group)) + geom_line() + scale_linetype_manual(values = c(1,2,3)) + geom_point(size=3) + scale_shape_manual(values = c(1,2,3)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) + scale_y_continuous(name = "mean(Weight) +/- se(Weight)")
```

SEE next step, here I used time > 1, but in this data the baseline was day 1.

```{r}
library(dplyr)
library(tidyr)
RaSuT <- RATSL %>% filter(Time > 0) %>% group_by(Group,ID) %>% summarise( mean=mean(Weight) ) %>% ungroup()
glimpse(RaSuT)
```


Looking for outliers without baseline (day 1)

```{r}
library(dplyr)
library(tidyr)
Rxx <- RATSL %>% filter(Time > 1) %>% group_by(Group,ID) %>% summarise( mean=mean(Weight) ) %>% ungroup()
glimpse(Rxx)
ggplot(Rxx, aes(x = Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") + scale_y_continuous(name = "mean(Weight), Time 8-64 days")
```

Seems that there are smaller outliers in groups 1 and 3, yet, in group 2 there is quite obvious outlier.



Filtering outliers
```{r}
RaSuTX <- Rxx %>% filter(mean < 550, mean > 250)
ggplot(RaSuTX, aes(x = Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") + scale_y_continuous(name = "mean(Weight), Time 8-64 days")
```
Now there is clearly a difference in group 2 boxplot. At this point group 3 outlier proved to be too difficult to exclude. Any tips to get it excluded?



Tests for differences

T-test, not possible because grouping factor has >2 levels. 
t.test(mean ~ Group, data = RaSuTX, var.equal = TRUE)

Anova (SEE before, RaSuT data has baseline with it).

```{r}
fit <- lm(mean ~ Group, data = RaSuT)
anova(fit)
```
Since P is very small the groups are different.




2) BPRS data analysis
```{r}
BPRSL <- read.table ("C:/Users/ville/OneDrive/Työpöytä/Open science kurssi 2020/IODS-project/Data/BPRSL.csv")
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)
dim(BPRSL); head(BPRSL); str(BPRSL)
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) + geom_line() + scale_linetype_manual(values = rep(1:10, times=4)) + facet_grid(. ~ treatment, labeller = label_both) + theme(legend.position = "none") + scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```


