---
title: "chapter6.Rmd"
output: html_document
---
# Ville Turppo, week 6



## 1) RATS data analysis swapped.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

RATSL <- read.table ("C:/Users/ville/OneDrive/Työpöytä/Open science kurssi 2020/IODS-project/Data/RATSL.csv")
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)
dim(RATSL); head(RATSL); str(RATSL)
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```
It seems that rats in group 1 are lighter than rats in the groups 2 and 3. In groups 2 and 3 the rats are heavier and gain more weight.

Standardized Weight
```{r}
RaS <- RATSL %>% group_by(Time) %>% mutate(stdweight = (Weight - mean(Weight))/sd(Weight) ) %>% ungroup()
glimpse(RaS)
ggplot(RaS, aes(x = Time, y = stdweight, group = ID)) + geom_line(aes(linetype = Group)) + scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) + scale_y_continuous(name = "Weight (standardized)") + theme(legend.position = "top")
```


Summary graphs

```{r}
n <- RaS$Time %>% unique() %>% length()
RaSu <- RaS %>% group_by(Group, Time) %>% summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>% ungroup()
ggplot(RaSu, aes(x = Time, y = mean, linetype = Group, shape = Group)) + geom_line() + scale_linetype_manual(values = c(1,2,3)) + geom_point(size=3) + scale_shape_manual(values = c(1,2,3)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) + scale_y_continuous(name = "mean(Weight) +/- se(Weight)")
```

SEE next step, here I used time > 1, but in this data the baseline was day 1.

```{r}
library(dplyr)
library(tidyr)
RaSuT <- RATSL %>% filter(Time > 0) %>% group_by(Group,ID) %>% summarise( mean=mean(Weight) ) %>% ungroup()
glimpse(RaSuT)
```


Looking for outliers without baseline (day 1)

```{r}
library(dplyr)
library(tidyr)
Rxx <- RATSL %>% filter(Time > 1) %>% group_by(Group,ID) %>% summarise( mean=mean(Weight) ) %>% ungroup()
glimpse(Rxx)
ggplot(Rxx, aes(x = Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") + scale_y_continuous(name = "mean(Weight), Time 8-64 days")
```

Seems that there are smaller outliers in groups 1 and 3, yet, in group 2 there is quite obvious outlier.



Filtering outliers
```{r}
RaSuTX <- Rxx %>% filter(mean < 550, mean > 250)
ggplot(RaSuTX, aes(x = Group, y = mean)) + geom_boxplot() + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") + scale_y_continuous(name = "mean(Weight), Time 8-64 days")
```
Now there is clearly a difference in group 2 boxplot. At this point group 3 outlier proved to be too difficult to exclude. Any tips to get it excluded?



Tests for differences

T-test, not possible because grouping factor has >2 levels. 
t.test(mean ~ Group, data = RaSuTX, var.equal = TRUE)

Anova (SEE before, RaSuT data has baseline with it).

```{r}
fit <- lm(mean ~ Group, data = RaSuT)
anova(fit)
```
Since P is very small the groups are different.




## 2) BPRS data analysis
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

BPRSL <- read.table ("C:/Users/ville/OneDrive/Työpöytä/Open science kurssi 2020/IODS-project/Data/BPRSL.csv")
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)
dim(BPRSL); head(BPRSL); str(BPRSL)
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) + geom_line() + scale_linetype_manual(values = rep(1:10, times=4)) + facet_grid(. ~ treatment, labeller = label_both) + theme(legend.position = "none") + scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```
There are 5 variables now in the data. In the plot it seems that both treatments work, yet, treatment 1 seem better at the end of follow-up.
Follow-up time was 8 weeks.



Creating linear multiple regression model of the BPRSL

```{r}
BPRSL_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRSL_reg)
```
This model assumes independence of the repeated measures of bprs, and this assumption is highly unlikely. And this model is not that suitable for this data.



Random intercept model
```{r}
library(lme4)
BPRSL_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref)
```


Slope model, week and subject as random effects
```{r}
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)

BPRSL_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref1)
anova(BPRSL_ref,BPRSL_ref1)
```


Random Intercept and Random Slope Model with interaction

```{r}
library(lme4)
library(dplyr)
library(tidyr)
library(ggplot2)

BPRSL_ref2 <- lmer(bprs ~ week * treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRSL_ref2)
anova(BPRSL_ref2,BPRSL_ref1)

Fitted <- fitted(BPRSL_ref2)
BPRSL <- BPRSL %>% mutate(Fitted)

ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))

ggplot(BPRSL, aes(x = week, y = Fitted, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$Fitted), max(BPRSL$Fitted)))
```

The AIC value in the summaries can be used to compare how well the models fit. Lower AIC = better. The AIC value is better when the greater amount of variation is explained by the model with the fewest possible variables. Using AIC I would say the last model with interaction is the best fitting model. BIC can also been used when choosing models. BIC penalizes model complexity (variables) more than AIC. BIC is lowest for the first model (2768). In all model summaries the is negative correation with weeks and bprs and also treatment and bprs = bprs scores decreasing during time and with treatments.
