---
title: "chapter3.Rmd"
output: html_document
---
# Week3 Analysis exercise, Ville Turppo 10.11.2020

Reading the data to R. Taking it from the internet incase I have made some mistakes before. Colnames command to see the variable names.

```{r}
AA <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt", sep = "," , header = TRUE)
colnames(AA)
```



Dim command shows that there are 382 rows = participants and 35 variables.
```{r}
dim (AA)
```


Logistic regression for factors relating to higher alcohol consumption.I choose to use sex, absences,failures and address.
My hypothesis: alco, absences and failures are related. Alco, sex and address are not related.

```{r}
m <- glm(high_use ~ failures + absences + sex + address, data = AA, family = "binomial")
```


Extracting coefficients and then intrepeting them ass Odds rations
And computing confidence intervals

OR <- coef(m) %>% exp
CI <- confint(m) %>% exp


Printing Odds Rations and confidence intrevals

cbind(OR, CI)


This is also my answer to question number 5 (logistic regression and its interpretation)
I would intepret the results as: Sex is the most related to high use, then failures, absences and adress = the further OR from zero, the more relation there is.

```{r}
library(tidyr)
library(dplyr); library(ggplot2)
```
Installed the packages

NEXT BOX-plots for
Intepretation of crosstabs and a boxplot. 
I only managed to make boxplot with highuse and absences, seem that those with high use have more absences. With crosstabs higher alco use doesn't mean more failures. Whereas men seem to have higher alco use and also students living in urban areas

```{r}
g1 <- ggplot(AA, aes(x = high_use, y = sex))
g2 <- ggplot(AA, aes(x = high_use, y = absences))
g3 <- ggplot(AA, aes(x = high_use, y = failures))
g4 <- ggplot(AA, aes(x = high_use, y = address))

g1 + geom_boxplot() + ggtitle("Student sex by alcohol consumption and sex")
g2 + geom_boxplot() + ggtitle("Student absences by alcohol consumption and sex")
g3 + geom_boxplot() + ggtitle("Student failing recent classes by alcohol consumption and sex")
g4 + geom_boxplot() + ggtitle("Student address by alcohol consumption and sex")

FAIL2 <- table(AA$failures, AA$high_use)
FAIL2
seex <- table(AA$sex, AA$high_use)
seex
add <- table(AA$address, AA$high_use)
add
```




New model for the predictions. Sex, absences and failures are inlcuded variables, since OR > 1.0.
```{r}
X <- glm(high_use ~ failures + absences + sex, data = AA, family = "binomial")
```

Predicting the probability of high_use
```{r}
probabilities <- predict(X, type = "response")
```

Adding the predicted probalities to data AA
```{r}
AA <- mutate(AA, probability = probabilities)
```

Using the probabilities to predict high_use
```{r}
AA <- mutate(AA, prediction = probability > 0.5)
```


Making Crosstabulation of the target variable vs predictions
```{r}
table(high_use = AA$high_use, prediction = AA$prediction)
```


Plotting the same results

```{r}
g <- ggplot(AA, aes(x = probability, y = high_use, col = prediction))
g + geom_point()
```



As seen in both, the crosstabs and the plot. Lot of FALSE predictions were infact TRUE. FALSE predictions were more accurate since only 12 real FALSEs were predicted as high use.
Sensitivity 26/112 * 100 = 23% and Specifity 258/270 *100 = 96%. Meaning the model is not good in predicting high alcohol use.


Bonus 10-fold cross-validation
Define a loss function (average prediction error)

```{r}
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}
```

The average number of wrong predictions in the data
```{r}
loss_func(class = AA$high_use, prob = AA$probability)
```



K-fold cross-validation
```{r}
library(boot)
CS <- cv.glm(data = AA, cost = loss_func, glmfit = X, K = 10)
```



Average number of wrong predictions in the cross validation
```{r}
CS$delta[1]
```

Seems that I'm doing something wrong, since not getting any answers


