---
title: "IODS course project"
author: "Ville Turppo"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_depth: 2
    fig_caption: true
    fig_width: 6
    fig_height: 4
---

***

# Introduction to Open Data Science - Course Project
Week 2 Data wrangling + Analysis

Wrangling

#opening the data.
test2 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt", sep="\t", header=TRUE)
dim(test2)
# row and columns
#checking data struckture
str(test2)
#str shows compactly the data

#Here making the combination variables.
deep_questions <- c("D03", "D11", "D19", "D27", "D07", "D14", "D22", "D30","D06",  "D15", "D23", "D31")
surface_questions <- c("SU02","SU10","SU18","SU26", "SU05","SU13","SU21","SU29","SU08","SU16","SU24","SU32")
strategic_questions <- c("ST01","ST09","ST17","ST25","ST04","ST12","ST20","ST28")
install.packages("dplyr")
library(dplyr)

deep_columns <- select(test2, one_of(deep_questions))
test2$deep <- rowMeans(deep_columns)

surface_columns <- select(test2, one_of(surface_questions))
test2$surf <- rowMeans(surface_columns)

strategic_columns <- select(test2, one_of(strategic_questions))
test2$stra <- rowMeans(strategic_columns)

# next keeping only 7 variables
columns <- c("gender","Age","Attitude", "deep", "stra", "surf", "Points")
end33 <- select(test2, one_of(columns))

colnames(end33)
#calculting the means.
end33$attitude <- end33$Attitude / 10
end33$straX <- end33$stra / 8
end33$deepX <- end33$deep / 12
end33$surfX <- end33$surf / 12
end33 <- filter(end33, Points > 0)

tolpat <- c("gender","Age","attitude", "deepX", "straX", "surfX", "Points")
final4 <- select(end33, one_of(tolpat))
dim(final4)

final4
#Saving my work
write.table(final4, file = "C:/Users/ville/OneDrive/Työpöytä/Open science kurssi 2020/IODS-project/data/week2.csv", sep = "\t")
#opening my work.
XX <- read.table("C:/Users/ville/OneDrive/Työpöytä/Open science kurssi 2020/IODS-project/data/week2.csv", sep = "\t")
head(XX)
str(XX)
dim(XX)


Analysis

# Here I'm opening the data from the internet. Sep tells R the seperator used in the data, header means there is a header.

AA <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", sep = ",", header = TRUE)
str(AA)
dim(AA)

# str command shows that the data has 7 variables and 166 rows.There are character(gender), integer(age/points) and numeric (the rest of variables) in the data.

library(ggplot2)
library(GGally)

# Library commands tell R to open libraries for the plot commands. First the ggplot2 and GGally packages need to be installed.

X <- ggpairs(AA)
X
D <- ggpairs(AA, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))
D

# this plot opens graphical overview of the data (e.g. distribution) and numerical values like correlation (i.e. how much the variables are related to one another. 0= no correlation, closer to 1 or -1 means the variables are more related)
# for example it seems that attitude is more related to points than age.

SS <- lm(points ~ attitude + stra + surf, data = AA)
summary(SS)
FF <- lm(points ~ attitude, data = AA)
summary(FF)

# I choose attitude, stra and surf as explanatory variables. P values for stra and surf had p>0.05 and thus low statistical relation with these variables and points. 

# These two were excluded and simple regression with points and attitude are shown below.

# fitted model: there seems to be statictically significant (low P value) relation between points and attitude.

# I learnt the following about R-squared from the internet today:
# R-squared is the percentage of the dependent variable variation that a linear model explains.
#100% represents a model that explains all of the variation in the response variable around its mean.
# Adjusted R-squared provides the same information as R-squared but adjusts for the number of terms in the model.
#plot command makes the following diagnostic plots and par command puts them on the same page.

par(mfrow = c(2,2))
plot(SS, which = c(1,2,5))
E <- lm(points ~ attitude + stra + surf, data = AA)
par(mfrow = c(2,2))
plot(E, which = c(1,2,5))

# Q-Q plot shows that both genders have same distribution.

# residuals Vs fitted: Residuals = Observed value – Fitted value. Y-axis= residual, X-axis= fitted values. Unwanted patterns of this plot = we can't trust the test. The should't be patterns. You shouldn't be able to predict the residual with fitted value, if the data is truly random.

# residuals vs leverage: can also be used to detect heteroskedasticity and non-linearity. Second, points with high leverage may be influential: that is, deleting them would change the model a lot. Points outside Cook's distance have high influence.

# Now there are few points outside Cook's distance, that could affect leverage and the model.




```{r child = "chapter1.Rmd"}
# This will include Chapter 1 (that is updated in its own file) in the document.
```
  
***
  
```{r child = "chapter2.Rmd"}
# This will include Chapter 2 (that is updated in its own file) in the document.
```

***

(more chapters to be added similarly as we proceed with the course!)

